---
title: "Portfolio 5- Study 2 data cleaning"
author: "Adam"
date: "5/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this markdown is to clean the data from Study 2 of my thesis.

## Set-up


```{r packages, echo=TRUE, message=FALSE}
library('reshape2')
library('dplyr')
library('Rcpp')
library('readr')
library('tidyverse')
library('psych')
```

### Pulling in the data

For this study, we have two datasets that we want to be able to combine into a single dataset. The first, Study 1 survey, has data from the survey we had participants take; and the second, Study 1 quiz, has their responses to the quiz they took.

First, we need to pull in the survey data

```{r bringing in survey data, echo=TRUE, collapse=TRUE, message=FALSE}
survey <- read_csv("data/Study 2 survey.csv")

# Ensuring the dataframe is a tibble.
as_tibble(survey)

#Taking a look at the table.
View(survey)
```

```{r bringing in survey data, echo=TRUE, collapse=TRUE, message=FALSE}
quiz <- read_csv("data/Study 2 Quiz.csv")

# Ensuring the dataframe is a tibble.
as_tibble(quiz)

#Taking a look at the table.
View(quiz)
```


Some cleaning was done in excel prior. 


As with Study 1, the data has already had its identifiers removed, but if it didn't, here is where that would be done. 

survey <- subset(survey, select= -c(StartDate, EndDate, Status,	IPAddress, Progress,
                Finished, RecordedDate, ResponseId, Durationinseconds, RecipientLastName, RecipientFirstName,
                RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude,
                DistributionChannel, UserLanguage))
                
The same was done to quiz before merging.

The other bit of cleaning I did was ensuring that participants input their participant id, which is used to identify them across the two datasets, in the same way. In actual practice, I have an excel document that tracks all the manual changes that were made in cleaning that up, but it is not linked here.

The next step will be to create the merged data file.

```{r data merging, echo=TRUE, collapse=TRUE, message=FALSE}
study1 <- survey %>% 
  full_join(quiz, by = "participant_id")

#Always make sure it's a tibble
as_tibble(study1)

View(study1)
```

### Descriptives cleaning

Fortunately, this is a great opportunity to ensure that the code from Portfolio 1 is working, since much of the cleaning is pretty similar. 

##### Gender dummy coding

First, I need to look at whether we have non-binary individuals in the dataset.

I did not change the variable names to the more convenient ones used in Study 1, so I am going to do that now.

```{r renaming gender variables, collapse=TRUE}
#male
study1 <- rename(study1, "demog_gender_male" = demog_gender_2)

#female
study1 <- rename(study1, "demog_gender_female" = demog_gender_1)

#trans
study1 <- rename(study1, "demog_gender_trans" = demog_gender_31)

#twospirit
study1 <- rename(study1, "demog_gender_twospirit" = demog_gender_32)

#genderqueer
study1 <- rename(study1, "demog_gender_genderqueer" = demog_gender_33)

#other
study1 <- rename(study1, "demog_gender_other" = demog_gender_34)

#no answer
study1 <- rename(study1, "demog_gender_no_answer" = demog_gender_4)
```

```{r checking gender demographics}
study1 %>%
        filter(is.na(demog_gender_female), is.na(demog_gender_male)) %>%
        count(demog_gender_trans, demog_gender_twospirit, demog_gender_genderqueer, demog_gender_other, demog_gender_no_answer)

study1 %>%
count(demog_gender_trans, demog_gender_twospirit, demog_gender_genderqueer, demog_gender_other, demog_gender_no_answer)

```

A single participant answered that they were genderqueer. This is too little to do anything with, so I will only dichotomize gender.

```{r Creating  binary gender dummy code}
study1 <- study1 %>%
        mutate(
        gender_male_dummy=case_when(
                demog_gender_male == 1 ~ 1,
                demog_gender_female == 1 ~ 0
                ))
```


##### Race dummy coding

The first time the code should be able to be pulled directly!

To create the race dummy codes, I first need to create a multiracial composite so that I can avoid adding multiracial people to individual dummy codes. 

Because the individual race codes have NA in all unfilled spots, I need to replace them with 0's to get the code to work correctly. The code is a little hideous and not consolidated, but I prefer to look and see them individually, especially because I can collapse this code in the chunk.

```{r adding 0 to original code}

study1 <- study1 %>%
        mutate(demog_race_51=case_when(
                demog_race_51 == 1 ~ 1,
                is.na(demog_race_51) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_52=case_when(
                demog_race_52 == 1 ~ 1,
                is.na(demog_race_52) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_59=case_when(
                demog_race_59 == 1 ~ 1,
                is.na(demog_race_59) ~ 0
                                       ))
study1 <- study1 %>%
        mutate(demog_race_21=case_when(
                demog_race_21 == 1 ~ 1,
                is.na(demog_race_21) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_22=case_when(
                demog_race_22 == 1 ~ 1,
                is.na(demog_race_22) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_23=case_when(
                demog_race_23 == 1 ~ 1,
                is.na(demog_race_23) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_29=case_when(
                demog_race_29 == 1 ~ 1,
                is.na(demog_race_29) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_31=case_when(
                demog_race_31 == 1 ~ 1,
                is.na(demog_race_31) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_32=case_when(
                demog_race_32 == 1 ~ 1,
                is.na(demog_race_32) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_33=case_when(
                demog_race_33 == 1 ~ 1,
                is.na(demog_race_33) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_39=case_when(
                demog_race_39 == 1 ~ 1,
                is.na(demog_race_39) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_11=case_when(
                demog_race_11 == 1 ~ 1,
                is.na(demog_race_11) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_12=case_when(
                demog_race_12 == 1 ~ 1,
                is.na(demog_race_12) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_13=case_when(
                demog_race_13 == 1 ~ 1,
                is.na(demog_race_13) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_19=case_when(
                demog_race_19 == 1 ~ 1,
                is.na(demog_race_19) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_41=case_when(
                demog_race_41 == 1 ~ 1,
                is.na(demog_race_41) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_42=case_when(
                demog_race_42 == 1 ~ 1,
                is.na(demog_race_42) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_89=case_when(
                demog_race_89 == 1 ~ 1,
                is.na(demog_race_89) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_99=case_when(
                demog_race_99 == 1 ~ 1,
                is.na(demog_race_99) ~ 0
                                       ))
```


With that task finished, we can work on making the multiracial composite.


```{r creating a multiracial sum}
study1 <- study1 %>%
        rowwise() %>%
        mutate(
                multiracial_sum=sum(c_across(demog_race_41:demog_race_89
        )))

```

```{r Creating multiracial dummy code}
study1 <- study1 %>%
        mutate(race_multiracial=case_when(
                multiracial_sum > 1 ~ 1,
                multiracial_sum == 1 ~ 0,
                multiracial_sum == 0 ~ 0
        ))

```


Now, I can make the racial composites. Due to sample size constraints I am going to have to make composites for several races. All composites will include their respective races and did not answer more than one race.

##### White composite

Composed of individuals who identified as European/European American, Middle Eastern/Middle Eastern American, Other White; or demog_race_51, demog_race_52, and demog_race_59.

```{r White composite dummy code}
study1 <- study1 %>%
        mutate(
        race_white=case_when(
                demog_race_51 | demog_race_52 | demog_race_59 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_white=case_when(
          race_white == 1 ~ 1,
          race_white == 0 ~ 0,
          is.na(race_white) ~ 0 
          ))

```

##### Black composite

Composed of individuals who identified as African American/Black, African, Caribbean, Other Black; or demog_race_21, demog_race_22, demog_race_23, and demog_race_29

```{r Black composite dummy code}
study1 <- study1 %>%
        mutate(
        race_black=case_when(
                demog_race_21 | demog_race_22 | demog_race_23 | demog_race_29 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_black=case_when(
          race_black == 1 ~ 1,
          race_black == 0 ~ 0,
          is.na(race_black) ~ 0 
          ))

```

##### Hispanic composite

Composed of individuals who identified as Mexican American/Chicano, Puerto Rican, Central American, Other Hispanic or Latino; or demog_race_31, demog_race_32, demog_race_33, and demog_race_39.

```{r Hispanic composite dummy code}
study1 <- study1 %>%
        mutate(
        race_hispanic=case_when(
                demog_race_31 | demog_race_32 | demog_race_33 | demog_race_39 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_hispanic=case_when(
          race_hispanic == 1 ~ 1,
          race_hispanic == 0 ~ 0,
          is.na(race_hispanic) ~ 0 
          ))

```

##### Asian composite

Composed of individuals who identified as East Asian (e.g., Chinese, Japanese, Korean, Taiwanese), Southeast Asian (e.g., Filipino, Vietnamese, Indonesian), South Asian (e.g., Pakistani, Indian, Nepalese, Sri Lankan), Other Asian; or demog_race_11, demog_race_12, demog_race_13, and demog_race_19.


```{r Asian composite dummy code}
study1 <- study1 %>%
        mutate(
        race_asian=case_when(
                demog_race_11 | demog_race_12 | demog_race_13 | demog_race_19 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_asian=case_when(
          race_asian == 1 ~ 1,
          race_asian == 0 ~ 0,
          is.na(race_asian) ~ 0 
          ))

```

##### Native composite

Composed of individuals who identified as Native Hawaiian or Other Pacific Islander, American Indian or Alaskan Native; or demog_race_41 and demog_race_42.


```{r Native composite dummy code}
study1 <- study1 %>%
        mutate(
        race_native=case_when(
                demog_race_41 | demog_race_42 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_native=case_when(
          race_native == 1 ~ 1,
          race_native == 0 ~ 0,
          is.na(race_native) ~ 0 
          ))

```

With these composites in hand, I can look more closely at how many we have in each category

```{r freq race}
table(study1$race_white)
table(study1$race_black)
table(study1$race_asian)
table(study1$race_hispanic)
table(study1$race_native)
table(study1$race_multiracial)
```

We have no native participants, so it is not needed. The majority of our sample is white, unsurprisingly.

#### Quiz recoding

We need to create dummy coded scores for the quiz, so I can get an average more easily.

```{r quiz question dummy codes}
#Question 1
study1 <- study1 %>%
        mutate(
        Quiz_Q1=case_when(
                Question_1 == 4 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q1=case_when(
          Quiz_Q1 == 1 ~ 1,
          is.na(Quiz_Q1) ~ 0 
          ))

# Question 2
study1 <- study1 %>%
        mutate(
        Quiz_Q2=case_when(
                Question_2 == 2 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q2=case_when(
          Quiz_Q2 == 1 ~ 1,
          is.na(Quiz_Q2) ~ 0 
          ))

# Question 3
study1 <- study1 %>%
        mutate(
        Quiz_Q3=case_when(
                Question_3 == 3 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q3=case_when(
          Quiz_Q3 == 1 ~ 1,
          is.na(Quiz_Q3) ~ 0 
          ))

# Question 4
study1 <- study1 %>%
        mutate(
        Quiz_Q4=case_when(
                Question_4 == 3 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q4=case_when(
          Quiz_Q4 == 1 ~ 1,
          is.na(Quiz_Q4) ~ 0 
          ))

# Question 5
study1 <- study1 %>%
        mutate(
        Quiz_Q5=case_when(
                Question_5 == 2 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q5=case_when(
          Quiz_Q5 == 1 ~ 1,
          is.na(Quiz_Q5) ~ 0 
          ))

#Question 6
study1 <- study1 %>%
        mutate(
        Quiz_Q6=case_when(
                Question_6 == 3 ~ 4
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q6=case_when(
          Quiz_Q6 == 1 ~ 1,
          is.na(Quiz_Q6) ~ 0 
          ))

# Question 7
study1 <- study1 %>%
        mutate(
        Quiz_Q7=case_when(
                Question_7 == 1 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q7=case_when(
          Quiz_Q7 == 1 ~ 1,
          is.na(Quiz_Q7) ~ 0 
          ))
#Question 8
study1 <- study1 %>%
        mutate(
        Quiz_Q8=case_when(
                Question_8 == 3 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q8=case_when(
          Quiz_Q8 == 1 ~ 1,
          is.na(Quiz_Q8) ~ 0 
          ))

#Question 9
study1 <- study1 %>%
        mutate(
        Quiz_Q9=case_when(
                Question_9 == 4 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q9=case_when(
          Quiz_Q9 == 1 ~ 1,
          is.na(Quiz_Q9) ~ 0 
          ))

```


Now that I have all the dummy coded versions, I can more easily compute the average score.

```{r quiz sum}
study1 <- study1 %>%
        rowwise() %>%
        mutate(
                quiz_average=mean(c_across(Quiz_Q1:Quiz_Q9
        )))

study1$quiz_average <- round(study1$quiz_average, digits = 2)

```

### Variables of Interest

Next, we need to create the composite variables for the variables of interest.


#### Manipulation check items

We had three manipulation check items.


##### Question disclosure

The first manipulation check item asked about how much the participants believed the questions asked them to disclose, consisting of two items: manip_personal and manip_encourage.

First, we should check the reliability.

```{r disclosure manip reliability}
alpha(study1[,c("manip_personal", "manip_encourage")])

```

Alpha is .73. Not amazing, but good enough to use.

Now, we can create the composite.

```{r disclosure manip composite}

study1$manip_check_comp2 <- rowMeans(study1[,c("manip_personal", "manip_encourage")], 
                              na.rm=TRUE)
```


##### Self-disclosure

The next manipulation check item asked about participants' actual levels of disclosure, consisting of three items: self_tell, self_share, and self_learn_about.

Checking the reliability.

```{r disclosure manip reliability}
alpha(study1[,c("self_tell", "self_share", "self_learn_about")])

```

Reliability is .89

Creating the composite.

```{r disclosure manip composite}

study1$self_disclosure_comp3 <- rowMeans(study1[,c("self_tell", "self_share", "self_learn_about")], 
                              na.rm=TRUE)
```


##### Group disclosure

The final manipulation check item asked about participants' beliefs about their group members' level of disclosure, consisting of three similar items to self-disclosure: group_tell, group_share, and group_learn_about.

Checking the reliability.

```{r disclosure manip reliability}
alpha(study1[,c("group_tell", "group_share", "group_learn_about")])

```

Reliability is .91

Creating the composite.

```{r disclosure manip composite}

study1$group_disclosure_comp3 <- rowMeans(study1[,c("group_tell", "group_share", "group_learn_about")], 
                              na.rm=TRUE)
```


##### Group talk

One question asked about whether groups talked, but that obvious doesn't need any cleaning here.



#### Institutional belonging

There are four variables that make up institutional belonging, "school_belong", "school_outsider", "school_comfort", and "school_fit_in".

One of the four variables needs to be reverse scored, school_outsider.

```{r reverse scoring WFU outsider}
study1 <- study1 %>%
        mutate(WFU_outsiderR= 8- WFU_outsider
                              )
```

Testing the reliability of the measure. Technically alpha gets masked by ggplot but I can just enable psych for a moment to override it.

```{r institutional belonging Reliability}
alpha(study1[,c("WFU_belong", "WFU_outsiderR", "WFU_comfort", "WFU_fit_in")])

```

Reliability is .92

Creating the actual composite.

```{r institutional belonging composite}

study1$school_belong_comp4 <- rowMeans(study1[,c("WFU_belong", "WFU_outsiderR", "WFU_comfort", "WFU_fit_in")], 
                              na.rm=TRUE)
```



#### Class belonging

Same as institutional belonging, there are the same four questions (with the prefix class_ this time) and I need to reverse score "class_outsider"

```{r reverse scoring class outsider}
study1 <- study1 %>%
        mutate(class_outsiderR= 8- class_outsider
        )

```

```{r class belonging reliability}
alpha(study1[,c("class_belong", "class_outsiderR", "class_comfort", "class_fit_in")])
```

Reliability is .89

```{r Class belonging composite}
study1$class_belong_comp4 <- rowMeans(study1[,c("class_belong", "class_outsiderR", "class_comfort", "class_fit_in")], 
                              na.rm=TRUE)
```



#### Closeness

We made a 3 item composite for how close participants felt to their group members, consisting of: self_close, self_like, and group_like.

```{r closeness reliability}
alpha(study1[,c("self_close", "self_like", "group_like")])
```

Reliability is .71

```{r closeness composite}
study1$close_comp3 <- rowMeans(study1[,c("self_close", "self_like", "group_like")], 
                              na.rm=TRUE)
```



#### Subject Area motivation

We asked about subject area motivation in two ways: intrinsic motivation and utility motivation.

##### Intrinsic motivation

The intrinsic motivation construct is made of two items: motiv_interesting and motiv_like.

```{r intrinsic motivation reliability}
alpha(study1[,c("motiv_interesting", "motiv_like")])
```

Reliability is .74

```{r intrinsic motivation composite}
study1$motiv_intrins_comp2 <- rowMeans(study1[,c("motiv_interesting", "motiv_like")], 
                              na.rm=TRUE)
```


##### Utility motivation

The utility motivation construct is made of three items: motiv_use_now, motiv_use_future, and motiv_use_grad

```{r utility motivation reliability}
alpha(study1[,c("motiv_use_now", "motiv_use_future", "motiv_use_grad")])
```

Reliability is .87

```{r utility motivation composite}
study1$motiv_util_comp3 <- rowMeans(study1[,c("motiv_use_now", "motiv_use_future", "motiv_use_grad")], 
                              na.rm=TRUE)
```



### Exploratory analyses

We had several questions that we were interested in on an exploratory basis.


#### Thinking deeper

One item asked about whether questions prompted participants to think deeper about the quiz topic.


#### Similarity to class

We asked participants about how similar the review task they did was to a task they think they would do in class. Two items measured this: manip_represent, manip_surprise. Surprise needs to be reverse-scored.

```{r reverse scoring surprise}
study1 <- study1 %>%
        mutate(manip_surpriseR= 6- manip_suprise
        )

```
Yes, there was a typo, but I don't want to go back and fix it since it already needs to be reverse scored.

```{r similarity reliability}
alpha(study1[,c("manip_represent", "manip_surpriseR")])
```

Reliability is .63. Which is really not great, but we did use it in my thesis.

```{r similarity composite}
study1$class_similar_comp2 <- rowMeans(study1[,c("manip_represent", "manip_surpriseR")], 
                              na.rm=TRUE)
```


#### Task enjoyment

We asked participants about how much they enjoyed the task, with three items: group_engage, group_enjoy, and group_fun.


```{r enjoyment reliability}
alpha(study1[,c("group_engage", "group_enjoy", "group_fun")])
```

Reliability is .87

```{r enjoyment composite}
study1$group_engagement_comp3 <- rowMeans(study1[,c("group_engage", "group_enjoy", "group_fun")], 
                              na.rm=TRUE)
```



### Saving data

```{r saving data}
write.csv(study1, file = "Study2_data_R_cleaned.csv")
```


It is finished!